A performance comparison of Dask and Spark for data-intensive neuroimaging pipelines

### 2019-07-25 latex/paper.tex
Typos

On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
slightly lower than Dask's in particular as the number of workers increase.
Within Dask, Delayed and Futures have a higher overhead than Bags. However,
overhead differences are compensated by an increase of data transfer time,
as a reduced overhead increases the concurrency between data transfers.

@@

Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when varying the
number of image blocks for a constant BigBrain image size. We were not able to run
Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size it can
compute. Once again, we do not observe any significant difference among the engines.
For all engines, makespan variability increases with the number of blocks, however,
engines scale very well in general.

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks the Dask Bag API has a much lower overhead time. This is because Dask
Bag was only using one thread per block in comparison to Dask Delayed and Dask
Futures which were offloading a block calculations on multiple threads of the same
worker. Once again, the data transfer time reduces with more blocks however the
overhead time increases by a similar amount. This is not observed for 30 blocks as
the workers are not used at full capacity, i.e., some threads are idle. Finally, the
variability of the overhead increases with the number of blocks, which explains the
makespan variability mentioned previously.

@@

%% Iterations
\subsection{Incrementation: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application while
varying the number of iterations. Overall,  Spark and Dask APIs are once
again equivalent, although Dask Delayed and Futures are slightly faster
than Bags and Spark for 1 and 10 few iterations, and Futures are faster than
Delayed, Bags and Spark for 100 iterations. Differences remain minor
though.

In Figure~\ref{fig:inc_tt_itr}, the total execution time breakdown is
shown. We observe good scalability of all the engines with the number of
iterations.
% TODO

@@

%% Sleep time
\subsection{Incrementation: Sleep delay}
Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation application
for different sleep delays. Overall, all engines again perform the same, and scale
well with task duration. Spark is initially slower than the Dask APIs, however, it is
faster when increasing the sleep delay. Also, within the Dask APIs, Dask Bag is
slower than the other two but it is not considerable.

@@

%%% HISTOGRAM %%%
\subsection{Histogram: Number of workers}
Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
amounts of workers. Spark is tremendiously faster than Dask APIs. The difference
narrows as the number of workers increases. Between Dask APIs there is no substantial
difference, however, Bag is slightly faster on average.

From Figure~\ref{fig:histo_tt_worker} the total time spend in each function is shown.
The computation time is significantly larger in Dask than Spark. This is potentially
due to Python's GIL preventing Dask to parallelize the computation on multiple threads.
Overall, the IO and overhead is comparable for all engines.

On Figure~\ref{fig:histo_tt_worker}, Dask engines have similar total execution time
when their number of workers varies, except for 8 worker where it incresase slightly.
On the other hand, Spark total execution time keeps increasing; especially at 8
workers. This is because Dask engines benefit more from additional workers.

@@

In figure~\ref{fig:histo_tt_block} the total time for each function is shown.
Note that at 30 blocks the worker's resources are not fully used since the workers
can process up to 64 task in parallel. Moreover, the overhead for 30 blocks is
erroneous due to a lower amount of block than available threads. This leads to thread
based engines to have higher overhead. Given that,

Spark has a much lower compute time which makes it significantly faster than Dask
engines. Overall, for all engines, lower block size results in lower IO time but
increased overhead time. Again, this is due to the overhead desynchronizing task thus
reducing the NFS bottle neck.
%TODO

@@

The reduce is done differently on the engines. While Spark and Delayed compute all
the reduce task on one process (thread) Spark computes it on the driver process. On
the other hand, Bag computes the reduce tasks on different workers and sends the
informations to the driver process afterwards.

@@

%%% HISTOGRAM (NumPy) %%%
\subsection{Histogram (NumPy): Number of workers}
Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount fo workers.
Overall, the makespan for all engines does not change considerably with an increase 
in worker. This is beacuse the engines do not scale well.

On figure~\ref{fig:histo_np_ms_worker}, there is no substantial difference between
 all engines, althought, Delayed is slightly faster.

From Figure~\ref{fig:histo_np_tt_worker}, the total execution time spent in each
function is shown. The overhead and read time increase proportionally to the number
of workers thus increasing the total time. This is because (1) most worker read at
the same time due to a low computation-IO ratio and (2) \MD{Determine why the
overhead is so large compared to the pure Python implementation}

@@

In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
Again, the overhead for 30 blocks is erroneous due a lower amount of task than
available thread which result in higher overhead for thread based engines.
Considering the overhead error and that at 30 blocks the worker only use half their
core, the total time is substantially the same for all block size. This is due lower
block sizes having a lower IO time, however, it is balanced by the resulting increase
in overhead.

@@

\subsection{Histogram (NumPy): Gantt chart}
Figure~\ref{fig:histo_np_gantt} shows the Gantt chart of the baseline Histogram
(NumPy) experiment. The overhead for Spark and Bag is mostly located between a read
and compute task or compute and read task. On the other hand, Delayed overhead is
dispersed between all type of tasks. This cause Delayed to overlap more compute and
read tasks amongs the different workers.

In figure~\ref{fig:histo_np_gantt}, The read tasks of all engines are similar initially, however, when computation starts
the subsequent read tasks are much shorter. This is because read tasks for Delayed
become desynchronized hence it significanlty reduces the IO bottle neck.

@@

In Figure~\ref{fig:bids_tt_worker}, the total execution time of each function is
shown. Overall, there is no significant difference between the engines. Each engines
have similar total time independently of their number of worker. This is because the
engines scale almost linearly. The overhead increase slowly as the number of workers
increases. This is due to more communication between the scheduler and workers as
well as inter-worker communication.

@@

%% Baseline gantt chart
\subsection{BidsApp example: Gantt chart}
Figure~\ref{fig:bids_gantt} shows the Gantt charts obtained for each engine and API.
The Gantt charts are structured in two main parts: participant analysis (orange) and
group analysis (blue). The participant analysis tasks differ greatly in lenght. This
is due to inqual amount of sessions per subject to process. The group analysis is
similar for all engines and APIs. Overall, most of the overhead encountered results
from the transition between the two analysis. This is because the group analysis is
requires the results of the participant analysis to start.

@@

Mathieu Dugr\'e was funded by a Undergraduate Student Research Assistant award from
the National Science and Engineering Research Council of Canada. We warmly thank
Compute Canada and \TG{Westgrid?} for providing the cloud infrastructure used in
these experiments, and the McGill Center for Integrative Neurocience for giving us
access to their cloud allocation.

>>>>

On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
slightly lower than Dask's in particular as the number of workers increase.
Within Dask, Delayed and Futures have a higher overhead than Bags. However,
overhead differences are compensated by an increase in data transfer time,
as a reduced overhead increases the concurrency between data transfers.

@@

Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when varying the
number of image blocks for constant BigBrain image size. We were not able to run
Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size it can
compute. Once again, we do not observe any significant difference among the engines.
For all engines, makespan variability increases with the number of blocks, however,
engines scale very well in general.

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks the Dask Bag API has a much lower overhead time. This is because Bag
was only using one thread per block in comparison to Delayed and Futures that were
offloading block calculations on multiple threads of the same worker. Once again,
the data transfer time reduces with more blocks however the overhead time increases
by a similar amount. This is not observed for 30 blocks as the workers are not used
at full capacity, i.e., some threads are idle. Finally, the variability of the
overhead increases with the number of blocks, which explains the makespan variability
mentioned previously.

@@

%% Iterations
\subsection{Incrementation: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application while
varying the number of iterations. Overall,  Spark and Dask APIs are once
again equivalent, although Dask Delayed and Futures are slightly faster
than Bags and, Spark for 1 and 10 few iterations, and Futures are faster than
Delayed, Bags and Spark for 100 iterations. Differences remain minor
though.

In Figure~\ref{fig:inc_tt_itr}, the total execution time breakdown is
shown. We observe the good scalability of all the engines with the number of
iterations.
% TODO

@@

%% Sleep time
\subsection{Incrementation: Sleep delay}
Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation application
for different sleep delays. Overall, all engines again perform the same and scale
well with task duration. Spark is initially slower than the Dask APIs, however, it is
faster when increasing the sleep delay. Also, within the Dask APIs, Dask Bag is
slower than the other two but it is not considerable.

@@

%%% HISTOGRAM %%%
\subsection{Histogram: Number of workers}
Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
amounts of workers. Spark is tremendously faster than Dask APIs. The difference
narrows as the number of workers increases. Between Dask APIs there is no substantial
difference, however, Bag is slightly faster on average.

From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is shown.
The computation time is significantly larger in Dask than Spark. This is potentially
due to Python's GIL preventing Dask to parallelize the computation on multiple threads.
Overall, the IO and overhead are comparable for all engines.

On Figure~\ref{fig:histo_tt_worker}, Dask engines have similar total execution time
when their number of workers varies, except for 8 workers where it increases slightly.
On the other hand, Spark total execution time keeps increasing; especially at 8
workers. This is because Dask engines benefit more from additional workers.

@@

In figure~\ref{fig:histo_tt_block} the total time for each function is shown.
Note that at 30 blocks the worker's resources are not fully used since the workers
can process up to 64 tasks in parallel. Moreover, the overhead for 30 blocks is
erroneous due to a lower amount of block than available threads. This leads to thread
based engines to have higher overhead. Given that,

Spark has a much lower compute time which makes it significantly faster than Dask
engines. Overall, for all engines, lower block size results in lower IO time but
increases the overhead time. Again, this is due to the overhead desynchronizing task
thus reducing the NFS bottleneck.
%TODO

@@

The reduce is done differently on the engines. While Spark and Delayed compute all
the reduce tasks on one process (thread) Spark computes it on the driver process. On
the other hand, Bag computes the reduce tasks on different workers and sends the
information to the driver process afterward.

@@

%%% HISTOGRAM (NumPy) %%%
\subsection{Histogram (NumPy): Number of workers}
Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount of workers.
Overall, the makespan for all engines does not change considerably with an increase
in the number of workers. This is because the engines do not scale well.

On figure~\ref{fig:histo_np_ms_worker}, there is no substantial difference between
 all engines, although, Delayed is slightly faster.

From Figure~\ref{fig:histo_np_tt_worker}, the total execution time spent in each
function is shown. The overhead and read time increase proportionally to the number
of workers thus increasing the total time. This is because (1) most workers read at
the same time due to a low computation-IO ratio and (2) \MD{Determine why the
overhead is so large compared to the pure Python implementation}

@@

In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
Again, the overhead for 30 blocks is erroneous due to a lower amount of task than
available threads which result in higher overhead for thread-based engines.
Considering the overhead error and that at 30 blocks the workers only use half their
core, the total time is substantially the same for all block size. This is due to
lower block sizes having a lower IO time, however, it is balanced by the resulting
increase in overhead.

@@

\subsection{Histogram (NumPy): Gantt chart}
Figure~\ref{fig:histo_np_gantt} shows the Gantt chart of the baseline Histogram
(NumPy) experiment. The overhead for Spark and Bag is mostly located between a read
and compute task or compute and read task. On the other hand, Delayed overhead is
dispersed between all types of tasks. This causes Delayed to overlap more compute and
read tasks among the different workers.

In figure~\ref{fig:histo_np_gantt}, The read tasks of all engines are similar initially, however, when computation starts
the subsequent read tasks are much shorter. This is because read tasks for Delayed
become desynchronized hence it significantly reduces the IO bottleneck.

@@

In Figure~\ref{fig:bids_tt_worker}, the total execution time of each function is
shown. Overall, there is no significant difference between the engines. Each engine
has a similar total time independently of its number of workers. This is because of
the, almost, linearly scale of the engines. The overhead increases slowly as the
number of workers increases. This is due to more communication between the scheduler
and workers as well as inter-worker communication.

@@

%% Baseline gantt chart
\subsection{BidsApp example: Gantt chart}
Figure~\ref{fig:bids_gantt} shows the Gantt charts obtained for each engine and API.
The Gantt charts are structured in two main parts: participant analysis (orange) and
group analysis (blue). The participant analysis tasks differ greatly in length. This
is due to the unequal amount of sessions per subject to process. The group analysis
is similar for all engines and APIs. Overall, most of the overhead encountered
results from the transition between the two analysis. This is because the group
analysis requires the results of the participant analysis to start.

@@

Mathieu Dugr\'e was funded by an Undergraduate Student Research Assistant award from
the National Science and Engineering Research Council of Canada. We warmly thank
Compute Canada and \TG{Westgrid?} for providing the cloud infrastructure used in
these experiments, and the McGill Center for Integrative Neuroscience for giving us
access to their cloud allocation.
